---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/README-"
)
```

[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis-CI Build Status](https://travis-ci.org/poissonconsulting/mcmcr.svg?branch=master)](https://travis-ci.org/poissonconsulting/mcmcr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/poissonconsulting/mcmcr?branch=master&svg=true)](https://ci.appveyor.com/project/poissonconsulting/mcmcr)
[![Coverage Status](https://img.shields.io/codecov/c/github/poissonconsulting/mcmcr/master.svg)](https://codecov.io/github/poissonconsulting/mcmcr?branch=master)
[![License: MIT](https://img.shields.io/badge/License-MIT-green.svg)](https://opensource.org/licenses/MIT)

# mcmcr

## Introduction

`mcmcr` is an R package to manipulate Monte Carlo Markov Chain (MCMC) samples.

For the purposes of this discussion, an MCMC sample represents the value of a term from a single iteration of a single chain (of a single analysis).
And while a simple parameter such as an intercept corresponds to a single term, more complex parameters such as an interaction between two factors consist of multiple terms with their own inherent dimensionality - in this case a matrix.
Not surprisingly a set of MCMC samples can be stored in different ways.
The three most common S3 classes store MCMC samples as follows:

- `coda::mcmc` stores the MCMC samples from a single chain as a matrix where each each row represents an iteration and each column represents a variable
- `coda::mcmc.list` stores multiple `mcmc` objects (with identical dimensions) as a list where each object represents a parallel chain
- `rjags::mcarray` stores the samples from a single parameter where the initial dimensions are the parameter dimensions, the second to last dimension is iterations and the last dimension is chains.

In the first two cases the terms/parameters are represented by a single dimension which means that the dimensionality inherent in the parameters is stored in the labelling of the variables, ie, `"bIntercept", "bInteraction[1,2]", "bInteraction[2,1]", ...`.
The structure of the `mcmc` and `mcmc.list` objects emphasizes the time-series nature of MCMC samples and is optimized for thining.
In contrast `mcarray` objects preserve the dimensionality of the parameters.

The `mcmcr` packages introduces two related S3 classes which also preserve the dimensionality of
the parameters:

- `mcmcr::mcmcarray` is very similar to `rjags::mcarray` except that the first dimension is the chains, the second dimension is iterations and the subsequent dimensions represent the dimensionality of the parameter (it is called `mcmcarray` to emphasize that the MCMC dimensions ie the chains and iterations come first);
- `mcmcr::mcmcr` stores multiple uniquely named `mcmcarray` objects with the same number of chains and iterations.

## Why mcmcr?

`mcmcarray` objects were introduced to facilitate manipulation and querying of the MCMC samples - although they are are just one `aperm` away from `mcarray` objects they are more intuitive to program with (at least for this programmer!)
`mcmcr` objects were introduced to allow a set of parameters to be manipulated and queried as a coherent whole.

The `mcmcr` package introduces a variety of (often) generic functions to manipulate and query `mcmcarray` and `mcmcr` objects. 
In particular it provides functions to

- coerce from and to `mcarray`, `mcmc` and `mcmc.list` objects;
- extract an objects `coef` table as a tibble;
- query an object's `nchains`, `niters`, `npars`, `nterms`, `nsims` and `nsams` as well as it's parameter dimensions (`pdims`)
- `subset` objects by chains, iterations and/or parameters;
- `bind_xx` a pair of objects by their chains, iterations, parameters or parameter dimensions;
- `combine_samples` of two objects by summing or otherwise aggregating their MCMC sample values;
- `collapse_chains` or `split_chains` an object's chains;
- `mcmc_map` over an objects values; 
- assess if an object has `converged` using `rhat` and `esr` (effectively sampling rate);
- and of course `thin`, `rhat`, `ess` (effective sample size), `print`, `plot` etc said objects.

## Demonstration

```{r}
library(mcmcr)

mcmcr_example

coef(mcmcr_example)
rhat(mcmcr_example, by = "term")
plot(mcmcr_example[["alpha"]])
```

## Installation

To install the latest version from GitHub
```
# install.packages("devtools")
devtools::install_github("poissonconsulting/mcmcr")
```

## Citation

```{r, comment="", echo=FALSE}
citation(package = "mcmcr")
```

## Contribution

Please report any [issues](https://github.com/poissonconsulting/mcmcr/issues).

[Pull requests](https://github.com/poissonconsulting/mcmcr/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

## Inspiration

[coda](https://github.com/cran/coda) and [rjags](https://github.com/cran/rjags)
