---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mcmcr

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![R-CMD-check](https://github.com/poissonconsulting/mcmcr/workflows/R-CMD-check/badge.svg)](https://github.com/poissonconsulting/mcmcr/actions)
[![Codecov test coverage](https://codecov.io/gh/poissonconsulting/mcmcr/branch/master/graph/badge.svg)](https://codecov.io/gh/poissonconsulting/mcmcr?branch=master)
[![License: MIT](https://img.shields.io/badge/License-MIT-green.svg)](https://opensource.org/licenses/MIT)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/mcmcr)](https://cran.r-project.org/package=mcmcr)
![CRAN Downloads](http://cranlogs.r-pkg.org/badges/mcmcr)
<!-- badges: end -->

`mcmcr` is an R package to manipulate Monte Carlo Markov Chain (MCMC) samples (Brooks et al. 2011).

## Installation

To install the latest release from [CRAN](https://cran.r-project.org)
```{r, eval=FALSE}
install.packages("mcmcr")
```

To install the developmental version from [GitHub](https://github.com/poissonconsulting/mcmcr)
```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("poissonconsulting/mcmcr")
```

## Introduction

For the purposes of this discussion, an MCMC *sample* represents the value of a *term* from a single *iteration* of a single *chain*.
While a simple *parameter* such as an intercept corresponds to a single term, more complex parameters such as an interaction between two factors consists of multiple terms with their own inherent dimensionality - in this case a matrix.
A set of MCMC samples can be stored in different ways.

### Existing Classes

The three most common S3 classes store MCMC samples as follows:

- `coda::mcmc` stores the MCMC samples from a single chain as a matrix where each each row represents an iteration and each column represents a variable
- `coda::mcmc.list` stores multiple `mcmc` objects (with identical dimensions) as a list where each object represents a parallel chain
- `rjags::mcarray` stores the samples from a single parameter where the initial dimensions are the parameter dimensions, the second to last dimension is iterations and the last dimension is chains.

In the first two cases the terms/parameters are represented by a single dimension which means that the dimensionality inherent in the parameters is stored in the labelling of the variables, ie, `"bIntercept", "bInteraction[1,2]", "bInteraction[2,1]", ...`.
The structure of the `mcmc` and `mcmc.list` objects emphasizes the time-series nature of MCMC samples and is optimized for thining.
In contrast `mcarray` objects preserve the dimensionality of the parameters.

### New Classes

The `mcmcr` package defines three related S3 classes which also preserve the dimensionality of
the parameters:

- `mcmcr::mcmcarray` is very similar to `rjags::mcarray` except that the first dimension is the chains, the second dimension is iterations and the subsequent dimensions represent the dimensionality of the parameter (it is called `mcmcarray` to emphasize that the MCMC dimensions ie the chains and iterations come first);
- `mcmcr::mcmcr` stores multiple uniquely named `mcmcarray` objects with the same number of chains and iterations.
- `mcmcr::mcmcrs` stores multiple `mcmcr` objects with the same parameters, chains and iterations.

All five classes (`mcmc`, `mcmc.list`, `mcarray`, `mcmcarray`, `mcmcr` and `mcmcrs`) are collectively referred to as MCMC objects.

## Why mcmcr?

`mcmcarray` objects were developed to facilitate manipulation of the MCMC samples.
`mcmcr` objects were developed to allow a set of dimensionality preserving parameters from a single analysis to be manipulated as a whole.
`mcmcrs` objects were developed to allow the results of multiple analyses using the same model to be manipulated together.

The `mcmcr` package (together with the [term](https://github.com/poissonconsulting/term) and [nlist](https://github.com/poissonconsulting/nlist) packages) introduces a variety of (often) generic functions to manipulate and query `mcmcarray`, `mcmcr` and `mcmcrs` objects (and `term` and `nlist` and `nlists` objects). 

In particular it provides functions to

- coerce from and to `mcarray`, `mcmc` and `mcmc.list` objects;
- extract an objects `coef` table (as a tibble);
- query an object's `nchains`, `niters`, `term::npars`, `term::nterms`, `nlist::nsims` and `nlist::nsams` as well as it's parameter dimensions (`term::pdims`) and term indices (`term::tindex`);
- `subset` objects by chains, iterations and/or parameters;
- `bind_xx` a pair of objects by their `xx_chains`, `xx_iterations`, `xx_parameters` or (parameter) `xx_dimensions`;
- combine the samples of two (or more) MCMC objects using `combine_samples` (or `combine_samples_n`) or combine the samples of a single MCMC object by reducing its dimensions using `combine_dimensions`;
- `collapse_chains` or `split_chains` an object's chains;
- `mcmc_map` over an objects values; 
- transpose an objects parameter dimensions using `mcmc_aperm`; 
- assess if an object has `converged` using `rhat` and `esr` (effectively sampling rate);
- and of course `thin`, `rhat`, `ess` (effective sample size), `print`, `plot` etc said objects.

The code is opinionated which has the advantage of providing a small set of stream-lined functions. 
For example the only 'convergence' metric is the uncorrected, untransformed, univariate split R-hat (potential scale reduction factor).
If you can convince me that additional features are important I will add them or accept a pull request (see below).
Alternatively you might want to use the `mcmcr` package to manipulate your samples before coercing them to an `mcmc.list` to take advantage of all the summary functions in packages such as `coda`.

## Demonstration

```{r}
library(mcmcr)

mcmcr_example

coef(mcmcr_example)
rhat(mcmcr_example, by = "term")
plot(mcmcr_example[["alpha"]])
```

## Inspiration

[coda](https://github.com/cran/coda) and [rjags](https://github.com/cran/rjags)

## Contribution

Please report any [issues](https://github.com/poissonconsulting/mcmcr/issues).

[Pull requests](https://github.com/poissonconsulting/mcmcr/pulls) are always welcome.

## Code of Conduct

Please note that the mcmcr project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.

## References

Brooks, S., Gelman, A., Jones, G.L., and Meng, X.-L. (Editors). 2011. Handbook for Markov Chain Monte Carlo. Taylor & Francis, Boca Raton. ISBN: 978-1-4200-7941-8.
