---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/README-"
)
```

![stability-unstable](https://img.shields.io/badge/stability-unstable-yellow.svg)
[![Travis-CI Build Status](https://travis-ci.org/poissonconsulting/mcmcr.svg?branch=master)](https://travis-ci.org/poissonconsulting/mcmcr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/poissonconsulting/mcmcr?branch=master&svg=true)](https://ci.appveyor.com/project/poissonconsulting/mcmcr)
[![Coverage Status](https://img.shields.io/codecov/c/github/poissonconsulting/mcmcr/master.svg)](https://codecov.io/github/poissonconsulting/mcmcr?branch=master)
[![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg)](https://opensource.org/licenses/MIT)

# mcmcr

## Introduction

`mcmcr` is an R package to manipulate Monte Carlo Markov Chain samples.

## Utilisation

```{r}
library(mcmcr)

mcmcr <- mcmcr:::mcmcr

nchains(mcmcr)
niters(mcmcr)
parameters(mcmcr)
nterms(mcmcr)

coef(mcmcr)
rhat(mcmcr)

estimates(mcmcr)

plot(mcmcr[[1]])
```

## Installation

To install the latest version from GitHub
```
# install.packages("devtools")
devtools::install_github("poissonconsulting/mcmcr")
```

## Contribution

Please report any [issues](https://github.com/poissonconsulting/mcmcr/issues).

[Pull requests](https://github.com/poissonconsulting/mcmcr/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

## Inspiration

- coda and rjags
